Nutritional Deficits and the Quantity-Quality Trade-off: Evidence from an Exogenous Fertility Shock in Low-Income Urban Settings in the Philippines

### 2025-06-10 docs/paper.Rmd
override AER default to display SCHUPF as issue name

---
title: "Nutritional Deficits and the Quantity-Quality Trade-off: Evidence from an Exogenous Fertility Shock in Low-Income Urban Settings in the Philippines"
short: "The Beckerian Quantity-Quality Trade-off in the Philippine Context"
journal: "2025 Schupf Fellowship Project" # AER, AEJ, PP, JEL
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
vol: 1
issue: 1
jel:
  - J13  # Fertility; Childbearing; Family Planning; Population Policy
  - I15  # Health and Economic Development
  - O15  # Human Resources; Human Development; Income Distribution; Migration
keywords:
  - fertility shocks
  - child nutrition
  - quantity-quality trade-off
  - urban poverty
  - Philippines
author:
  - name: Erika Salvador
    firstname: Erika
    surname: Salvador
    email: esalvador28@amherst.edu
    affiliation: Amherst College
  - name: Caroline Theoharides
    firstname: Caroline
    surname: Theoharides
    email: ctheoharides@amherst.edu
    affiliation: Amherst College
acknowledgements: |
  This research was made possible by the Schupf Fellowship at Amherst College. I am indebted to my faculty adviser, Professor Caroline Theoharides, for her mentorship and support throughout the development of this project. I also wish to thank Faculty Director Professor Amelie Hastie and the campus partners whose institutional support enabled this work. I am further grateful to the Departments of Economics and Mathematics & Statistics for their academic support. The views expressed and any errors contained herein are entirely my own responsibility.
abstract: |
  This paper examines whether increased fertility affects early-life nutritional outcomes in low-income urban households. I exploit a natural experiment created by a 1990 policy in Manila, Philippines, which banned modern contraceptives from city-run health facilities. Using a difference-in-differences framework and nationally representative data from the Philippine Demographic and Health Surveys, I estimate the reduced-form impact of the policy on child height-for-age and weight-for-height. [Will add more after data analysis]
output:
  rticles::aea_article:
    keep_tex: true
    citation_package: natbib
bibliography: references.bib
---

>>>>

---
title: "Nutritional Deficits and the Quantity-Quality Trade-off: Evidence from an Exogenous Fertility Shock in Low-Income Urban Settings in the Philippines"
short: "The Beckerian Quantity-Quality Trade-off in the Philippine Context"
journal: "" # AER, AEJ, PP, JEL
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
vol: 1
issue: 1
jel:
  - J13  # Fertility; Childbearing; Family Planning; Population Policy
  - I15  # Health and Economic Development
  - O15  # Human Resources; Human Development; Income Distribution; Migration
keywords:
  - fertility shocks
  - child nutrition
  - quantity-quality trade-off
  - urban poverty
  - Philippines
author:
  - name: Erika Salvador
    firstname: Erika
    surname: Salvador
    email: esalvador28@amherst.edu
    affiliation: Amherst College
  - name: Caroline Theoharides
    firstname: Caroline
    surname: Theoharides
    email: ctheoharides@amherst.edu
    affiliation: Amherst College
acknowledgements: |
  This research was made possible by the Schupf Fellowship at Amherst College. I am indebted to my faculty adviser, Professor Caroline Theoharides, for her mentorship and support throughout the development of this project. I also wish to thank Faculty Director Professor Amelie Hastie and the campus partners whose institutional support enabled this work. I am further grateful to the Departments of Economics and Mathematics & Statistics for their academic support. The views expressed and any errors contained herein are entirely my own responsibility.
abstract: |
  This paper examines whether increased fertility affects early-life nutritional outcomes in low-income urban households. I exploit a natural experiment created by a 1990 policy in Manila, Philippines, which banned modern contraceptives from city-run health facilities. Using a difference-in-differences framework and nationally representative data from the Philippine Demographic and Health Surveys, I estimate the reduced-form impact of the policy on child height-for-age and weight-for-height. [Will add more after data analysis]
output: rticles::aea_article
bibliography: references.bib
---
